# Dimensionality reduction techniques

```{r}
library(ggplot2)
library(GGally)
library(dplyr)
library(corrplot)
```

## Overview of the data

```{r} 
human <- read.table("human.txt", header = TRUE)
summary(human)
str(human)
dim(human)

```

The data contains eight variables and 155 rows. 

The dataset is a combined from a data of Human Development index and Gender Inequality index. These datasets are from United Nations Development Programme’s Human Development reports.

See the datapage for further information [here][id].

[http://hdr.undp.org/en/content/human-development-index-hdi][id]

Technical notes can be found [here][id].

[http://hdr.undp.org/sites/default/files/hdr2015_technical_notes.pdf][id]



```{r}
ggpairs(human)
```


```{r}
cor(human) %>% corrplot
```

## PCA on non-standardized data

```{r}
# Mean again zero. Check how affects the variance
# perform principal component analysis (with the SVD method)
pca_human <- prcomp(human)

biplot(pca_human, choices = 1:2, cex = c(0.6, 1), col = c("grey40", "deeppink2"))
```


## Standardize variables

```{r}
human_std <- scale(human)

pca_human <- prcomp(human_std)



# rounded percentages of variance captured by each PC
pca_pr <- round(100*s$importance[2,], digits = 1) 

# print out the percentages of variance
pca_pr

# create object pc_lab to be used as axis labels
pc_lab <- paste0(names(pca_pr), " (", pca_pr, "%)")

# draw a biplot
biplot(pca_human, cex = c(0.8, 1), col = c("grey40", "deeppink2"), xlab = pc_lab[1], ylab = pc_lab[2])

```

```{r}

```

## Interpertation

## Tea drinkers

```{r}
library(FactoMineR)
library(ggplot2)
library(tidyr)
library(stringr)
data(tea)

# column names to keep in the dataset
keep_columns <- c("Tea", "How", "how", "sugar", "where", "lunch")

# select the 'keep_columns' to create a new dataset
tea_time <- select(tea, one_of(keep_columns))

# look at the summaries and structure of the data
summary(tea_time)
str(tea_time)

# visualize the dataset
gather(tea_time) %>% ggplot(aes(value)) + facet_wrap("key", scales = "free") + geom_bar() + theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8))
summary(tea_time)
dim(tea_time)
str(tea_time)
```

```{r}
# multiple correspondence analysis
mca <- MCA(tea_time, graph = FALSE)

# summary of the model
summary(mca)

# visualize MCA
plot(mca, invisible=c("ind"), habillage = "quali")
```



```{r}
plotellipses(mca, keepvar=(1:5))
```

